---
title: "Fangchao_LysisComparisib"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


this was done in UNIX 


Do I need to uzip raw files for cutadapt ?  I think I did.
gunzip *


ls -1 | grep "R1" > files_R1files
awk ' FS="." {print$1"_cut."$2} '  files_R1files  >  files_R1files_cut
paste files_R1files_cut files_R1files  > files_R1_command
vi files_R1_command
%s/^/\~\/.local\/bin\/cutadapt -g GGCCTACGGGAGGCAGCAG -o /g





ls -1 | grep "R2" > files_R2files
awk ' FS="." {print$1"_cut."$2} '  files_R2files  >  files_R2files_cut
paste files_R2files_cut files_R2files  > files_R2_command
vi files_R2_command

%s/^/\~\/.local\/bin\/cutadapt -g CCGGACTACHVGGGTWTCTAAT -o /g.  # this was meesed up a bit and I didn't get the ~ 
chmod +x files_R1_command


Librarys to load
========================================================
```{r  packagestoload, cache=TRUE }
library(reshape2)
library(ggplot2)
library(qiimer)
library(vegan)
library(shape)
library(RColorBrewer)
library(plyr)
library(GUniFrac)
library(ape)
library(dendextend)
library(ade4)
library(geiger)
library(gtools)
library(phytools)
library(devtools)
library("Biostrings")
library(DESeq2)     #see notes about install above
library(phyloseq)  #had to install differently
library(dada2)
library(cowplot) # caused a weird dependency issue
library(data.table)  #caused a weird dependency issue
library(ampvis2)  #had to install differently
library(ShortRead) #see notes about install above
library(backports)
library(dada2)
library(genefilter)


```



```{r  loading data files,  cache = TRUE}
# path = "/Users/jenniferkuehl/Documents/Projects/On_going/Amplicon/Fangchao_lysismethod/cut_files"
path = "cut_file"
fnFs <- sort(list.files(path, pattern="R1_001_cut.fastq", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="R2_001_cut.fastq", full.names = TRUE))
sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)
filtFs <- file.path(path, "filtered", paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(path, "filtered", paste0(sample.names, "_R_filt.fastq.gz"))

```





```{r    filter read denoise, cache=TRUE}
out_Fsonly <- filterAndTrim(fnFs, filtFs, truncLen=300,
              maxN=0, maxEE=3, truncQ=2, 
              compress=TRUE, multithread=TRUE )  
not.lost <- file.exists(filtFs)
filtFs <- filtFs[not.lost]   

errF <- learnErrors(filtFs, multithread=TRUE)
derepFs <- derepFastq(filtFs, verbose=TRUE) 
dadaFs <- dada(derepFs, err=errF, multithread=TRUE) 
```


Start here.
```{r  maketable and assigntax , cache=TRUE}
Fonly = makeSequenceTable(dadaFs)
Fonly.nochim <- removeBimeraDenovo(Fonly, method="consensus", multithread=TRUE, verbose=TRUE)
sum(Fonly.nochim)/sum(Fonly)

getN <- function(x) sum(getUniques(x))
track_F <- cbind(out_Fsonly, sapply(dadaFs, getN), rowSums(Fonly.nochim))
colnames(track_F) <- c("input", "filtered", "denoisedF", "nonchim")
rownames(track_F) <- sample.names
head(track_F)



taxa2 <- assignTaxonomy(Fonly.nochim, "silva_nr_v132_train_set.fa.gz", multithread=TRUE)
taxa2 <- addSpecies(taxa2, "silva_species_assignment_v132.fa.gz")
```



```{r}
#because the ESv are names by sequence,  not always convenient,  except when merging dataframes.

taxa3 = taxa2
taxa3 = as.data.frame(taxa3)
taxa3$Sequence = rownames(taxa3)
taxa3 = taxa3[,c(8,1,2,3,4,5,6,7)]
rownames(taxa3) = NULL
newnames = paste0("esv", 1:length(rownames(taxa3)))
rownames(taxa3) = newnames
taxa4 = as.matrix(taxa3)

colnames(Fonly.nochim) = newnames

samples2.out <- rownames(Fonly.nochim)
subject <- sapply(strsplit(samples2.out, "_"), `[`, 1)
rownames(Fonly.nochim) = subject


```

```{r  phyloseq time ,  cache=TRUE}
meta.txt = read.csv(file="P21P23MetaData20190829.csv", header=TRUE,stringsAsFactors=FALSE)
row.names(meta.txt) = meta.txt$Sample_ID

ps <- phyloseq(otu_table(Fonly.nochim, taxa_are_rows=FALSE), 
               sample_data(meta.txt), 
               tax_table(taxa4))
ps.t = transform_sample_counts(ps, function(OTU) OTU/sum(OTU))   #2045 taxa
flist<- filterfun(kOverA(5, 1e-04))
ps.t.filtered = filter_taxa(ps.t, flist ,TRUE)  #194

```



```{r}
tax_export = tax_table(ps)
otu_export = otu_table(ps)
sample_export = sample_data(ps)
write.table(tax_export, file = "tax_export", sep = "\t",  quote = FALSE)
write.table(otu_export, file = "otu_export", sep = "\t",  quote = FALSE)
write.table(sample_export, file = "sample_export", sep = "\t",  quote = FALSE)

```





```{r   for ampvis graphs}

otutable <- data.frame(OTU = colnames(phyloseq::otu_table(ps)@.Data),
                       phyloseq::t(otu_table(ps))@.Data,    #transform since taxa were not in rows
                       phyloseq::tax_table(ps)@.Data,
                       check.names = FALSE
                       )

metadata <- data.frame(phyloseq::sample_data(ps), 
                       check.names = FALSE
                       )

otunames= names(otutable)
amp<- amp_load(otutable, metadata)
```
